---
title: "Loan Classification and Credit Score Framework"
author: "Kenneth I. D., Ayooluwa Adedipe, Patience Kadiri"
date: "3/6/2020"
output: html_document
---



library(plyr)
library(dplyr)
library(tidyr)
library(tidyverse)
library(data.table)
library(caret)
library(mlbench)
library(logitnorm)
library(mlr)


Importing the all_accounts dataset

```{r, echo=FALSE}
library(plyr)
accounts <- fread('all_accounts.csv')

count(accounts, CCY)
```
Converting all currencies to Naira

```
newCCY <- accounts %>%
  mutate(CCY=recode(CCY, `NGN`=1, `GBP`=471.9, `USD`= 365.5, `EUR`=398.33))
```

Setting customer id as a string

```{r, echo=FALSE}
accounts$CUSTOMER_UNIQUE_ID <- sprintf("%f", accounts$CUSTOMER_UNIQUE_ID)
newCCY$CUSTOMER_UNIQUE_ID <- sprintf("%f", newCCY$CUSTOMER_UNIQUE_ID)
```

Multiplying the balance and the new CCY in the newCCY table

```{r, echo=FALSE}
newAVL_BAL <- as.data.frame(newCCY$ACY_AVL_BAL*newCCY$CCY)
newAVL_BAL$new_AVL_BAL <- newAVL_BAL$`newCCY$ACY_AVL_BAL * newCCY$CCY`
```

Dropping columns on the new table

```{r, echo=FALSE}
newAVL_BAL <- subset(newAVL_BAL, select = c(new_AVL_BAL))
```
Merging newAVL_BAL to the accounts table

```{r, echo=FALSE}
accounts <- c(accounts, newAVL_BAL)
accounts <- as.data.frame(accounts)
```

Dropping columns on the accounts table

```{r, echo=FALSE}
accounts <- subset(accounts, select = -c(ACY_AVL_BAL, CCY))
accounts <- subset(accounts, select = -c(RECORD_STAT, MASKED_ACCOUNT))
```

Sum the available balance

```{r, echo=FALSE}
availBalance <- accounts %>% 
  group_by(CUSTOMER_UNIQUE_ID) %>%
  summarise(AVAIL_BALANCE = sum(new_AVL_BAL))

availBalance$AVAIL_BALANCE <- availBalance$AVAIL_BALANCE
```

Assigning a rating scale to the balances

```{r, echo=FALSE}
availBalance$BALANCE_RATING = 
  ifelse(availBalance$`AVAIL_BALANCE` <= 1000, 4,
    ifelse(availBalance$`AVAIL_BALANCE` <= 5000, 5,
      ifelse(availBalance$`AVAIL_BALANCE` <= 10000, 6, 
        ifelse(availBalance$`AVAIL_BALANCE` <= 50000, 6,
          ifelse(availBalance$`AVAIL_BALANCE` <= 100000, 7,
            ifelse(availBalance$`AVAIL_BALANCE` <= 500000, 8,
              ifelse(availBalance$`AVAIL_BALANCE` <= 1000000, 9, 10)))))))

acctBalanceRating = subset(availBalance, select = -c(AVAIL_BALANCE))
```

```{r, echo=FALSE}
count(availBalance, BALANCE_RATING)
```

WE WILL MAKE USE OF THE ACCOUNT_BALANCE_RATING TABLE FOR MERGING. THIS IS THE AMOUNT OF MONEY AVAILABLE IN THE CUSTOMER'S ACCOUNT AS AT THE LAST BOOKING DATE ON THE DATASET.

Importing the all_balances dataset

```{r, echo=FALSE}
balances <- fread('all_balances.csv')
```

Make Customer Unique ID a string variable

```{r, echo=FALSE}
balances$CUSTOMER_UNIQUE_ID <- sprintf("%f", balances$CUSTOMER_UNIQUE_ID)
```

Summing up all the balances 

```{r, echo=FALSE}
allBalances <- balances %>%
  group_by(CUSTOMER_UNIQUE_ID) %>%
  summarize(VALUE_DATE=max(VAL_DT), HIGHEST_BALANCE= max(BAL), MEAN_BALANCE= mean(BAL), BALANCE_HISTORY= sum(BAL))

balanceHistory = subset(allBalances, select= -c(VALUE_DATE, MEAN_BALANCE))
```

GRADING THE HIGHEST BALANCE OF THE CUSTOMERS

```{r, echo=FALSE}
balanceHistory$HIGHEST_BALANCE_RATING = 
  ifelse(balanceHistory$`HIGHEST_BALANCE` <= 1000, 4,
    ifelse(balanceHistory$`HIGHEST_BALANCE` <= 5000, 5,
      ifelse(balanceHistory$`HIGHEST_BALANCE` <= 10000, 6, 
        ifelse(balanceHistory$`HIGHEST_BALANCE` <= 50000, 6,
          ifelse(balanceHistory$`HIGHEST_BALANCE` <= 100000, 7,
            ifelse(balanceHistory$`HIGHEST_BALANCE` <= 500000, 8,
              ifelse(balanceHistory$`HIGHEST_BALANCE` <= 1000000, 9, 10)))))))
```

Grading the BALANCE_HISTORY of the customers
```{r, echo=FALSE}
balanceHistory$BALANCE_HISTORY_RATING = 
  ifelse(balanceHistory$`BALANCE_HISTORY` <= 1000, 4,
    ifelse(balanceHistory$`BALANCE_HISTORY` <= 5000, 5,
      ifelse(balanceHistory$`BALANCE_HISTORY` <= 10000, 6, 
        ifelse(balanceHistory$`BALANCE_HISTORY` <= 50000, 6,
          ifelse(balanceHistory$`BALANCE_HISTORY` <= 100000, 7,
            ifelse(balanceHistory$`BALANCE_HISTORY` <= 500000, 8,
              ifelse(balanceHistory$`BALANCE_HISTORY` <= 1000000, 9, 10)))))))

count(balanceHistory, HIGHEST_BALANCE_RATING)
count(balanceHistory, BALANCE_HISTORY_RATING)

balanceHistoryRating <- subset(balanceHistory, select = -c(HIGHEST_BALANCE, BALANCE_HISTORY))
```

We make use of both the HIGHEST_BALANCE and the BALANCE_HISTORY table for our analysis. This table contains the highest account balance in the customers' histories. It also contains the sum of all the amounts that has passed through the customers' accounts

I DID NOT RUN THE FOLLOWING CODES BELOW BECAUSE OF THE TOLL THEY TAKE ON MY PC. INSTEAD, I TOOK THE PATIENCE OF RECODING THE DATASET, DROPPING SOME COLUMNS AND PARSING INTO A NEW CSV FILE 

Importing the all_transactions_digital table

__trans <- fread('all_transactions_digital.csv')__

setting customer id as a string
__trans$CUSTOMER_UNIQUE_ID <- sprintf("%f", trans$CUSTOMER_UNIQUE_ID)__
__newTrans <- subset(trans, select= c(CUSTOMER_UNIQUE_ID, AC_CCY, LCY_AMOUNT, DRCR_IND))__
__newTrans <- newTrans %>%mutate(AC_CCY=recode(AC_CCY, `NGN`=1, `GBP`=471.9, `USD`= 365.5, `EUR`=398.33))__

multiplying the LCY_AMOUNT and EX_RATE in the table
__newTrans$AMOUNT <- as.data.frame(newTrans$AC_CCY*newTrans$LCY_AMOUNT)__
__newTrans$AMOUNT <- AMOUNT.newTrans$AC_CCY*newTrans$LCY_AMOUNT__

Categorising the Product with Dummy Variables
library(dummies)
__transClean <- dummy(newTrans$DRCR_IND, sep="_")__
__transClean <- as.data.frame(transClean)__
__transClean$DRCR_IND <- transClean__ this puts all the new dummy columns back into the data frame
__newTrans$CREDIT <- transClean$DRCR_IND_C__
__newTrans$DEBIT <- transClean$DRCR_IND_D__
__newTrans$CUSTOMER_UNIQUE_ID = sprintf("%f", newTrans$CUSTOMER_UNIQUE_ID)__

__cashFlow <- newTrans %>%
group_by(CUSTOMER_UNIQUE_ID, CREDIT, DEBIT) %>%
summarize(TOTAL_CREDIT = CREDIT%*%LCY_AMOUNT, TOTAL_DEBIT = LCY_AMOUNT%*%DEBIT)
write.csv(cashFlow, file = 'cashFlow.csv')__

Importing the cashFlow dataset

```{r, echo=FALSE}
library(tidyverse)
cashFlow <- fread('cashFlow.csv')
```

Converting the Customer Unique ID to a string variable

```{r, echo=FALSE}
cashFlow$CUSTOMER_UNIQUE_ID = sprintf("%f", cashFlow$CUSTOMER_UNIQUE_ID)
```

Summing up the all the credits on the dataset per Customer Unique ID

```{r, echo=FALSE}
cashFlowHistory <- cashFlow %>%
  group_by(CUSTOMER_UNIQUE_ID) %>%
  summarize(TOTAL_CREDIT = sum(TOTAL_CREDIT), TOTAL_DEBIT = sum(TOTAL_DEBIT))

cashFlowHistory$Balance <- cashFlowHistory$TOTAL_CREDIT - cashFlowHistory$TOTAL_DEBIT
```

Grading the Customer's Credit

```{r, echo=FALSE}
cashFlowHistory$Balance_Rating = 
  ifelse(cashFlowHistory$`Balance` <= 0, 0,
    ifelse(cashFlowHistory$`Balance` <= 1000, 4,
      ifelse(cashFlowHistory$`Balance` <= 5000, 5,
        ifelse(cashFlowHistory$`Balance` <= 10000, 6, 
          ifelse(cashFlowHistory$`Balance` <= 50000, 6,
            ifelse(cashFlowHistory$`Balance` <= 100000, 7,
              ifelse(cashFlowHistory$`Balance` <= 500000, 8,
                ifelse(cashFlowHistory$`Balance` <= 1000000, 9, 10))))))))

count(cashFlowHistory, Balance_Rating)
```

DROPPING COLUMNS

```{r, echo=FALSE}
cashFlowBalanceRating <- subset(cashFlowHistory, select = c(CUSTOMER_UNIQUE_ID, Balance_Rating))
```

This dataset contains the sum of both credit and debit transactions on each customer's account within the period specified. It also contains the net difference between them, but we will use the cashFlow$Balance_Rating for our credit score framework.

```{r, echo=FALSE}
bureau <- fread('bureau_score.csv')
count(bureau, bureau$BUREAU_SCORE)
count(bureau, bureau$BAL_NON_DELQ)
count(bureau, bureau$BAL_30DPD)
colnames(bureau)
```

Customer unique id as a string

```{r, echo=FALSE}
bureau$CUSTOMER_UNIQUE_ID <- sprintf("%f", bureau$CUSTOMER_UNIQUE_ID)

debt_history <- bureau%>%
  group_by(CUSTOMER_UNIQUE_ID) %>%
  summarise(total30DPD=sum(BAL_30DPD), total60DPD=sum(BAL_60DPD), total90DPD=sum(BAL_PL_90DPD))

debt_history$total_debt <- debt_history$total30DPD + debt_history$total60DPD + debt_history$total90DPD

debt_history$Debt_Rating = 
  ifelse(debt_history$`total_debt` <= 0, 10,
    ifelse(debt_history$`total_debt` <= 1000, 9,
      ifelse(debt_history$`total_debt` <= 5000, 8,
        ifelse(debt_history$`total_debt` <= 10000, 7, 
          ifelse(debt_history$`total_debt` <= 50000, 6,
            ifelse(debt_history$`total_debt` <= 100000, 5,
              ifelse(debt_history$`total_debt` <= 500000, 3,
                ifelse(debt_history$`total_debt` <= 1000000, 1, 0))))))))

count(debt_history, Debt_Rating)
unique(debt_history)
```

Dropping columns

```{r, echo=FALSE}
Debt_Profile <- subset(debt_history, select = c(CUSTOMER_UNIQUE_ID, Debt_Rating))
```

WE WILL MAKE USE OF THE DEBT_PROFILE TABLE FOR FURTHER ANALYSIS

```{r, echo=FALSE}
invest <- fread('all_investments.csv')
count(invest, PRODUCT)
count(invest, invest$PRODUCT_NAME)
count(invest, invest$ROLLOVER_ALLOWED)
count(invest, invest$ROLLOVER_COUNT)
count(invest, invest$CONTRACT_STATUS)
colnames(invest)
```

Setting customer id as a string

```{r, echo=FALSE}
invest$CUSTOMER_UNIQUE_ID <- sprintf("%f", invest$CUSTOMER_UNIQUE_ID)
```

Dropping columns

```{r, echo=FALSE}
newInvest = subset(invest, select = -c(CONTRACT_REF_NO, PAYMENT_METHOD, CURRENCY, AMOUNT, ROLLOVER_ALLOWED, ROLLOVER_COUNT, CONTRACT_STATUS, MASKED_ACCOUNT))
```

Performing a Split-Apply-Combine to find the Investment strength of the Customer

```{r, echo=FALSE}
investHistory <- newInvest %>%
  group_by(CUSTOMER_UNIQUE_ID) %>% 
  summarize(BOOKING_DATE=max(BOOKING_DATE), MATURITY_DATE=max(MATURITY_DATE), COMBINED_TENOR= sum(TENOR), MAIN_COMP_RATE=max(MAIN_COMP_RATE), INVESTMENT =sum(LCY_AMOUNT))

investHistory$R.O.INVESTMENT <- investHistory$MAIN_COMP_RATE * investHistory$INVESTMENT
```

Dropping Columns 

```{r, echo=FALSE}
investmentHistory <- subset(investHistory, select = -c(BOOKING_DATE, MATURITY_DATE, MAIN_COMP_RATE))
```

RATING THE INVESTMENTS AND THE ROI OF THE CUSTOMERS

```{r, echo=FALSE}
investmentHistory$Investment_Rating = 
  ifelse(investmentHistory$`INVESTMENT` <= 0, 0,
    ifelse(investmentHistory$`INVESTMENT` <= 1000, 2,
      ifelse(investmentHistory$`INVESTMENT` <= 5000, 3,
        ifelse(investmentHistory$`INVESTMENT` <= 10000, 4, 
          ifelse(investmentHistory$`INVESTMENT` <= 50000, 5,
            ifelse(investmentHistory$`INVESTMENT` <= 100000, 7,
              ifelse(investmentHistory$`INVESTMENT` <= 500000, 8,
                ifelse(investmentHistory$`INVESTMENT` <= 1000000, 9, 10))))))))
```

RATING THE ROI OF THE CUSTOMERS

```{r, echo=FALSE}
investmentHistory$ROI_Rating = 
  ifelse(investmentHistory$`R.O.INVESTMENT` <= 0, 10000,
    ifelse(investmentHistory$`R.O.INVESTMENT` <= 100000, 4,
      ifelse(investmentHistory$`R.O.INVESTMENT` <= 500000, 5,
        ifelse(investmentHistory$`R.O.INVESTMENT` <= 1000000, 6, 
          ifelse(investmentHistory$`R.O.INVESTMENT` <= 5000000, 6,
            ifelse(investmentHistory$`R.O.INVESTMENT` <= 10000000, 7,
              ifelse(investmentHistory$`R.O.INVESTMENT` <= 50000000, 8,
                ifelse(investmentHistory$`R.O.INVESTMENT` <= 100000000, 9, 10))))))))

Investment_Profile <- subset(investmentHistory, select =c(CUSTOMER_UNIQUE_ID, Investment_Rating, ROI_Rating))
```

This contains the amount of investment a customer has in the bank. It is equal to the investment strength of the customer. We will make use of the Investment Profile table for further analysis.

Importing the all_loans table

```{r, echo=FALSE}
loans <- fread('all_loans.csv')
count(loans, loans$CURRENCY)
count(loans, loans$ACCOUNT_STATUS)
```

Finding the duration of the loan

```{r, echo=FALSE}
TENURE <- as.Date.character((loans$MATURITY_DATE), format="%m/%d/%Y")- as.Date.character((loans$BOOK_DATE), format="%m/%d/%Y")
loans$TENURE <- TENURE
```

Recoding the Product code

```{r, echo=FALSE}
library(dummies)
loansClean <- dummy(loans$PRODUCT_CODE, sep="_")
loansClean <- as.data.frame(loansClean)
loansClean <- subset(loansClean, select= c(PRODUCT_CODE_AMPC, PRODUCT_CODE_PDLP))
#this puts all the new dummy columns back into the data frame
loans$PRODUCT_CODE <- loansClean
loans <- c(loans, loansClean)
loans <- as.data.frame(loans)
```

Renaming the new columns

Making customer ID a character variable

```{r, echo=FALSE}
loans$CUSTOMER_UNIQUE_ID <- sprintf("%f", loans$CUSTOMER_UNIQUE_ID)
```

Summing up tenure to get the loan history by customer id

```{r, echo=FALSE}
HISTORY <-
  aggregate(loans$TENURE, by=list(CUSTOMER_UNIQUE_ID=loans$CUSTOMER_UNIQUE_ID), FUN=sum )

HISTORY$history <- HISTORY$x
HISTORY <- subset(HISTORY, select= -c(x))
```

Summing up amount the amount financed to get the total amount of loans each customer has recieved

```{r, echo=FALSE}
loanHistory <-
  aggregate(loans$AMOUNT_FINANCED, by=list(CUSTOMER_UNIQUE_ID=loans$CUSTOMER_UNIQUE_ID), FUN=sum )
loanHistory$loanHistory <- loanHistory$x
loanHistory<- subset(loanHistory, select = -c(x))
```

Merging the history and loan history tables as a scoring table

```{r, echo=FALSE}
Credit <- merge(HISTORY, loanHistory, by=intersect(names(HISTORY), names(loanHistory)))
```

Removing colums

```{r, echo=FALSE}
newLoans <- subset(loans, select = -c(LOAN_REF, PRODUCT_CODE, PRODUCT_NAME, BOOK_DATE, MATURITY_DATE, CURRENCY, ACCOUNT_STATUS))
```

Rating the history(days) and loanHistory (total loan amount) per customer

```{r, echo=FALSE}
Credit$history_Rating =
  ifelse(Credit$`history` <= 300, 4,
    ifelse(Credit$`history` <= 500, 5,
      ifelse(Credit$`history` <= 1000, 6, 
        ifelse(Credit$`history` <= 2000, 6,
          ifelse(Credit$`history` <= 3000, 7,
            ifelse(Credit$`history` <= 4000, 8,
              ifelse(Credit$`history` <= 5000, 9, 10)))))))

Credit$loanHistory_Rating =
  ifelse(Credit$`loanHistory` <= 3000, 3,
    ifelse(Credit$`loanHistory` <= 10000, 4,
      ifelse(Credit$`loanHistory` <= 100000, 5, 
        ifelse(Credit$`loanHistory` <= 1000000, 6,
          ifelse(Credit$`loanHistory` <= 10000000, 7,
            ifelse(Credit$`loanHistory` <= 100000000, 8,
              ifelse(Credit$`loanHistory` <= 900000000, 9, 10)))))))

CreditProfile <- subset(Credit, select =  c(CUSTOMER_UNIQUE_ID, history_Rating, loanHistory_Rating))
```

We will use the Credit Profile table for further analysis

```{r, echo=FALSE}
loansApp <- fread('all_loans_application.csv')
count(loansApp, APPLICATION_STATUS)
unique(loansApp)
count(loansApp, MODULE)
count(loansApp, DATE_OF_APPLY)
count(loansApp, CUSTOMER_UNIQUE_ID)
```

Setting Customer ID as a character variable

```{r, echo=FALSE}
loansApp$CUSTOMER_UNIQUE_ID <- sprintf("%f", loansApp$CUSTOMER_UNIQUE_ID)
```

Dropping missing values, diong a summary statistics on the dataset and grading the loan application of the customers

```{r, echo=FALSE}
library(tidyr)
loansApp %>% drop_na()
#aggregating the date of application per customer
appDate <- loansApp %>%
  group_by(CUSTOMER_UNIQUE_ID, MODULE, APPLICATION_STATUS, LOAN_CLASSIFICATION) %>%
  summarize(DATE_OF_APPLICATION = max(DATE_OF_APPLY), MEAN_APPLIED_AMOUNT = mean(APPLIED_AMOUNT), MEAN_INCR_ALLOWED= mean(INCR_ALLOWED))

LoanAppCredit <- subset(appDate, select = c(CUSTOMER_UNIQUE_ID, MEAN_APPLIED_AMOUNT, LOAN_CLASSIFICATION))

LoanAppCredit$NEW_LOAN_RATING = 
  ifelse(LoanAppCredit$`MEAN_APPLIED_AMOUNT` <= 5000, 5,
    ifelse(LoanAppCredit$`MEAN_APPLIED_AMOUNT` <= 10000, 6, 
      ifelse(LoanAppCredit$`MEAN_APPLIED_AMOUNT` <= 50000, 7,
        ifelse(LoanAppCredit$MEAN_APPLIED_AMOUNT <= 100000, 8,
          ifelse(LoanAppCredit$MEAN_APPLIED_AMOUNT <= 500000, 8,
            ifelse(LoanAppCredit$MEAN_APPLIED_AMOUNT <= 1000000, 9, 10))))))
```


Importing the demographics dataset

```{r, echo=FALSE}
demo <- fread(file='all_demographics.csv', header = TRUE, sep = "|", stringsAsFactors = FALSE)
```

Selecting columns, and recoding the columns that I need for further analysis

```{r, echo=FALSE}
newDemo <- subset(demo, select = c(CUSTOMER_UNIQUE_ID, CUSTOMER_CATEGORY, MINOR, MARITAL_STATUS, DEPENDENT_OTHERS, SEX))

demographics <- newDemo

demographics <- demographics %>% mutate(CUSTOMER_CATEGORY=recode(CUSTOMER_CATEGORY, `C`= 10, `CORP`=10, `GOVT`=10, `HNI`=10, `I`=5, `IND`=7, `MINORS`=1, `RSA`=10, `S`=5, `SAL`=7, `STUDENTS`=4, `SUPERMARKT`=7, `TIER1`=10, `TIER2`=8, `WALKIN`=3))

demographics <- demographics %>% mutate(MINOR=recode(MINOR, `N`= 10, `Y`=1, `M`=1, `NA `=1))

demographics <- demographics %>% mutate(MARITAL_STATUS=recode(MARITAL_STATUS, `S`= 10, `P`=5, `M`=8, `E`=5, `D`=5, `NA`=1))

demographics <- demographics %>% mutate(DEPENDENT_OTHERS=recode(DEPENDENT_OTHERS, `0`= 10, `NA`=1))

demographics <- demographics %>% mutate(SEX=recode(SEX, `M`=7 , `F`=10, `P`=1, `NA `=1))

demographics <- demographics %>% mutate_all(~replace(.,is.na(.), 1))

```


MERGING ALL THE DATA SETS TOGETHER

```{r, echo=FALSE}
df <-merge(acctBalanceRating, balanceHistoryRating, by=intersect(names(acctBalanceRating), names(balanceHistoryRating)), by.acctBalanceRating ="CUSTOMER_UNIQUE_ID", by.balanceHistoryRating ="CUSTOMER_UNIQUE_ID", all=TRUE, all.acctBalanceRating=all, all.balanceHistoryRating=all, sort=TRUE, suffixes=c(".acctBalanceRating", ".balanceHistoryRating"), no.dups=TRUE, incomparables = NULL)

df <-merge(df, CreditProfile, by=intersect(names(df), names(CreditProfile)), by.df ="CUSTOMER_UNIQUE_ID", by.CreditProfile ="CUSTOMER_UNIQUE_ID", all=TRUE, all.df=all, all.CreditProfile=all, sort=TRUE, suffixes=c(".df", ".CreditProfile"), no.dups=TRUE, incomparables = NULL)

df <-merge(df, cashFlowBalanceRating, by=intersect(names(df), names(cashFlowBalanceRating)), by.df ="CUSTOMER_UNIQUE_ID", by.cashFlowBalanceRating ="CUSTOMER_UNIQUE_ID", all=TRUE, all.df=all, all.cashFlowBalanceRating=all, sort=TRUE, suffixes=c(".df", ".cashFlowBalanceRating"), no.dups=TRUE, incomparables = NULL)

df <-merge(df, Debt_Profile, by=intersect(names(df), names(Debt_Profile)), by.df ="CUSTOMER_UNIQUE_ID", by.Debt_Profile ="CUSTOMER_UNIQUE_ID", all=TRUE, all.df=all, all.Debt_Profile=all, sort=TRUE, suffixes=c(".df", ".Debt_Profile"), no.dups=TRUE, incomparables = NULL)

df <-merge(df, demographics, by=intersect(names(df), names(demographics)), by.df ="CUSTOMER_UNIQUE_ID", by.demographics ="CUSTOMER_UNIQUE_ID", all=TRUE, all.df=all, all.demographics=all, sort=TRUE, suffixes=c(".df", ".demographics"), no.dups=TRUE, incomparables = NULL)

df <-merge(df, LoanAppCredit, by=intersect(names(df), names(LoanAppCredit)), by.df ="CUSTOMER_UNIQUE_ID", by.LoanAppCredit ="CUSTOMER_UNIQUE_ID", all=TRUE, all.df=all, all.LoanAppCredit=all, sort=TRUE, suffixes=c(".df", ".LoanAppCredit"), no.dups=TRUE, incomparables = NULL)
```

Assigning all NA values to 1. This is to be on the safe side as the customers affected do not have these information on their profile. Rather than completely discard them, which will ultimately have a negative effect on our data set, we set the NA values to the least rating (just to assume worse case scenerios for the customers whose details are missing, resulting in the NA values). When the customers eventually provide these information when they come to apply for loans, their credit scores can then be boosted appropriately.

Recoding the Loan_Classification column

```{r, echo=FALSE}
df <- df %>% mutate(LOAN_CLASSIFICATION=recode(LOAN_CLASSIFICATION, `PERFORMING`= 1, `NON-PERFORMING`=0))

df <- df %>% mutate_all(~replace(.,is.na(.), 1))
df <- as.numeric(df)


count(df, df$BALANCE_RATING)
count(df, df$HIGHEST_BALANCE_RATING)
count(df, df$BALANCE_HISTORY_RATING)
count(df, df$history_Rating)
count(df, df$loanHistory_Rating)
count(df, df$Balance_Rating)
count(df, df$Debt_Rating)
unique(df)
count(df, df$CUSTOMER_CATEGORY)
count(df, df$MARITAL_STATUS)
count(df, df$LOAN_CLASSIFICATION)
```

Dropping columns

```{r, echo=FALSE}
df <- subset(df, select = -c(MEAN_APPLIED_AMOUNT))
```

Summing up all rows

```{r, echo=FALSE}
df$TOTAL_SCORE <- df$BALANCE_RATING + df$HIGHEST_BALANCE_RATING + df$BALANCE_HISTORY_RATING + df$history_Rating + df$loanHistory_Rating + df$Balance_Rating + df$Debt_Rating + df$CUSTOMER_CATEGORY + df$MINOR + df$MARITAL_STATUS + df$DEPENDENT_OTHERS + df$SEX + df$NEW_LOAN_RATING

# Assigning the TOTAL POSSIBLE SCORE
df$TOTAL_POSSIBLE_SCORE <- 130

# Finding the PERCENTAGE of the CREDIT_SCORES
df$CREDIT_SCORE_PERCENT <- ((df$TOTAL_SCORE/df$TOTAL_POSSIBLE_SCORE)*100)
df$CREDIT_SCORE_PERCENT <-round(df$CREDIT_SCORE_PERCENT)

# Assigning the MINIMUM_PERCENT of SCORES allowed
df$MINIMUM_PERCENT_SCORE <- 50

# Indexing the CUSTOMER_UNIQUE_ID to become a row name, as a means of identifying the customers rather than a variable in the data set

count <- count(df, df$CREDIT_SCORE_PERCENT)
max(df$TOTAL_SCORE)

df$CUSTOMER_UNIQUE_ID <- as.numeric(df$CUSTOMER_UNIQUE_ID)
```


Regression Analysis for my credit score model

First dividing the data set into train and test set...

```{r, echo=FALSE}
dt = sort(sample(nrow(df), nrow(df)*.7))
train <- (df[dt,])
test <- (df[-dt,])

# Search for redundant variables

set.seed(1000000000)
CorrelationMatrix <- cor(df)
library(corrplot)
corrplot::corrplot(CorrelationMatrix,  method = 'square') 
corrplot(CorrelationMatrix, type = "lower")  
corrplot.mixed(CorrelationMatrix, lower.col = "black", number.cex = .9)
```


Train the model: 

Setting the cross validation parameters, Recursive Feature Elimination using the Boruta library

```{r, echo=FALSE}
library(randomForest)
library(Boruta)

boruta.train <- Boruta(CREDIT_SCORE_PERCENT~ train$BALANCE_RATING+train$HIGHEST_BALANCE_RATING+train$BALANCE_HISTORY_RATING+train$history_Rating+train$loanHistory_Rating+train$Balance_Rating+train$Debt_Rating+train$CUSTOMER_CATEGORY+train$MINOR+train$MARITAL_STATUS+train$DEPENDENT_OTHERS+train$SEX+train$NEW_LOAN_RATING, data=train)

print(boruta.train)

plot(boruta.train, xlab = "", xaxt= "n")
```

The blue boxplots corresponds to minimal credit scores while the blue boxplots corresponds to tentative and confirmed attributes. Doing the final Boruta

```{r, echo=FALSE}
final.boruta <- TentativeRoughFix(boruta.train)
print(final.boruta)
# No attribute or variable is deemed unfit for our model. So we proceed to the final result.
getSelectedAttributes(final.boruta, withTentative = F)

#Creating a dataframe to store up the final boruta
df_boruta <- attStats(final.boruta)
class(df_boruta)
print(df_boruta)
```

All 13 variables are important in our model, so we use all 13 to create a model for the credit score framework.

```{r, echo=FALSE}
CSF <- lm(CREDIT_SCORE_PERCENT~ train$BALANCE_RATING +train$HIGHEST_BALANCE_RATING +train$BALANCE_HISTORY_RATING +train$history_Rating +train$loanHistory_Rating +train$Balance_Rating+ train$Debt_Rating +train$CUSTOMER_CATEGORY +train$MINOR +train$MARITAL_STATUS +train$DEPENDENT_OTHERS +train$SEX +train$NEW_LOAN_RATING, data=train)

plot(CSF)
summary(CSF)
anova(CSF)
```

Predictive analysis for the credit score framework

```{r, echo=FALSE}
predict(CSF, data.frame(test))
predCSF <- as.data.frame(predict(CSF, data.frame(test)))
```




Loan Classification Model (Logit Model)

Recursive Feature Selection for the Loan Classification model using the Boruta library also.

```{r, echo=FALSE}
set.seed(12)
boruta2.train <- Boruta(LOAN_CLASSIFICATION~ BALANCE_RATING + HIGHEST_BALANCE_RATING + BALANCE_HISTORY_RATING + history_Rating + loanHistory_Rating + Balance_Rating + Debt_Rating + CUSTOMER_CATEGORY + MINOR + MARITAL_STATUS + DEPENDENT_OTHERS + SEX + NEW_LOAN_RATING, data=train)

print(boruta2.train)
plot(boruta2.train)
```

The blue boxplots corresponds to minimal and average variables, the black boxplot represents the rejected variables while the blue boxplots corresponds to tentative and confirmed attributes. Doing the final Boruta for this model

```{r, echo=FALSE}
last.boruta <- TentativeRoughFix(boruta2.train)
print(last.boruta)
# 5 attribute or variable is deemed unfit for our model. So we remove them from our model
getSelectedAttributes(last.boruta, withTentative = F)
```

Creating a dataframe to store up the last boruta

```{r, echo=FALSE}
df_last_boruta <- attStats(last.boruta)
class(df_last_boruta)
print(df_last_boruta)
```

The Model

```{r, echo=FALSE}
LCM <- glm(LOAN_CLASSIFICATION ~ BALANCE_RATING + HIGHEST_BALANCE_RATING + BALANCE_HISTORY_RATING + history_Rating + loanHistory_Rating + Balance_Rating + Debt_Rating + NEW_LOAN_RATING, data=train, family = binomial(link = "logit"))
summary(LCM)
```

This model can explain 83% of the variance of the dataset. Generally, loans are seen as non performing when their tenure has exceeded a period of 90days. All the variables in the models have a way of affecting the customer's ability to make sure their loans do not exceed the 90 days tenure.

Predictive analysis for the loan classification model

```{r, echo=FALSE}
predLCM <- predict(LCM, data.frame(test))
invlogit(predLCM)

table(test$LOAN_CLASSIFICATION, invlogit(predLCM) > 0.5)
```

37,949 observations are real Truths of Performing loans, 552 are False-Truths of Non-Performing loans, 91 observations are False Performing loans, while 83 are False Non Performing loans. 


Conclusion
